Software architectures to integrate workflow engines in science gateways

### 2016-04-25 paper.tex
edits to section 2

\begin{figure*}
\centering
\hspace*{0.2\columnwidth}
\subfloat[Tight integration]{
\def\svgwidth{0.6\columnwidth}
\input{figures/tight.pdf_tex}
\label{archi:tight}
}
\hfill
\subfloat[Service invocation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/service.pdf_tex}
\label{archi:service}
\hspace*{0.2\columnwidth}
}\\
\hspace*{0.2\columnwidth}
\subfloat[Sub-tasking]{
\def\svgwidth{0.6\columnwidth}
\input{figures/sub-task.pdf_tex}
\label{archi:sub-task}
}
\hfill
\subfloat[Pool]{
\def\svgwidth{0.6\columnwidth}
\input{figures/agent.pdf_tex}
\label{archi:agent}
\hspace*{0.2\columnwidth}
}\\
\subfloat[Nested workflows. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-1.pdf_tex}
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-2.pdf_tex}
\label{archi:nested}
}\\
\subfloat[Workflow import. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.5\columnwidth}
\input{figures/import-1.pdf_tex}
\hfill
\def\svgwidth{0.5\columnwidth}
\input{figures/import-2.pdf_tex}
\label{archi:import}
}
\caption{Architectures}
\label{fig:architectures}
\end{figure*}
\begin{table*}
\centering
\footnotesize
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{Systems} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{Ardizzone2012}, 
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, e-BioInfra~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{krüger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012}, 
            Science gateways in ~\cite{kacsuk2014science}$^w$
            } \\
  Sub-task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$\todo{others}\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$
\end{tabular}
\caption{Classification of science gateways based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the system
in~\cite{wu2010accelerating} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture where the
\texttt{d} interaction is implemented as a set of calls to Pegasus
command-line tools (e.g. pegasus-status, pegasus-statistics,
etc). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The pool model was implemented in the SHIWA pool~\cite{ROGE-13}
diagrammed in Figure~\ref{fig:shiwa-pool-architecture}. In this
Figure, interaction \texttt{d} of Figure~\subref*{archi:agent} is
implemented in 3 different types of calls to the pool: workflow
submission (\texttt{1} \& \texttt{2}), workflow status retrieval
(\texttt{a} and \texttt{b}), and workflow retrieval (\texttt{13} \&
\texttt{14}). In Figure ~\ref{fig:shiwa-pool-architecture},
interaction \texttt{f} of Figure~\subref*{archi:agent} is implemented
through 2 types of calls: workflow instance retrieval (\texttt{4},
\texttt{5} and \texttt{6}), and workflow instance update (\texttt{11}
and \texttt{12}). Workflow instance retrieval is used by the agents to
fetch work from the pool. Workflow instance update is used by the
agents to update workflow statuses. In the SHIWA pool, agents can wrap
different types of workflow engines to execute workflows expressed
with different languages. Calls \texttt{0}, \texttt{7}, \texttt{8} and
\texttt{9} on Figure~\ref{fig:shiwa-pool-architecture} are used by
workflow engine plugins to declare their supported language and launch
engines, and by workflow engines to report their status to the
agent. These calls are specific to the SHIWA Pool implementation of
the \texttt{agent} component and therefore have no corresponding
representation in Figure~\subref*{archi:agent}.

@@

Nested workflows are abstract architectural patterns that can be
instantiated in the various architectures described previously. We
focus on instantiation with the service invocation model
(Figure~\subref*{archi:nested}-Right) as this is the most used
architecture. In such an instantiation, we assume that the parent and
child workflow engines are distinct pieces of software that require
different workflow services invoked by distinct \texttt{d}
interactions. If this is not the case, then workflow services can be
collapsed in a single one with a \texttt{d} interaction with
itself. Workflow engines communicate with infrastructures using
\texttt{b} and \texttt{c$_2$}. Science gateway transfers user data to
infrastructures using \texttt{c$_1$} interactions.

@@

See Figure~\subref*{archi:import}. This is an abstract model that we
instantiate with the service invocation architecture for
consistency. Workflows are integrated in the science gateway through
format conversion from a native format to the science gateway
format. Workflow import is usually an offline process which is not
involved in workflow execution. Workflow import was implemented in the SHIWA
Simulation Platform through the IWIR language that provided a common
language for portability across grid workflow
systems~\cite{plankensteiner-prodan-etal:2013}.

>>>>

\begin{figure*}
\centering
\hspace*{0.2\columnwidth}
\subfloat[Tight integration]{
\def\svgwidth{0.6\columnwidth}
\input{figures/tight.pdf_tex}
\label{archi:tight}
}
\hfill
\subfloat[Service invocation]{
\def\svgwidth{0.6\columnwidth}
\input{figures/service.pdf_tex}
\label{archi:service}
\hspace*{0.2\columnwidth}
}\\
\hspace*{0.2\columnwidth}
\subfloat[Sub-tasking]{
\def\svgwidth{0.6\columnwidth}
\input{figures/sub-task.pdf_tex}
\label{archi:sub-task}
}
\hfill
\subfloat[Pool]{
\def\svgwidth{0.6\columnwidth}
\input{figures/agent.pdf_tex}
\label{archi:agent}
\hspace*{0.2\columnwidth}
}\\
\subfloat[Nested workflows. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-1.pdf_tex}
\def\svgwidth{0.9\columnwidth}
\input{figures/nested-2.pdf_tex}
\label{archi:nested}
}\\
\subfloat[Workflow import. Left: abstract model. Right: instantiation with service invocation.]{
\def\svgwidth{0.5\columnwidth}
\input{figures/import-1.pdf_tex}
\hfill
\def\svgwidth{0.5\columnwidth}
\input{figures/import-2.pdf_tex}
\label{archi:import}
}
\caption{Architectures}
\label{fig:architectures}
\end{figure*}
\begin{table*}
\centering
\footnotesize
\rowcolors{1}{white}{gray!25}
\begin{tabular}{ll}
  \textbf{Architecture} & \textbf{Systems} \\
  \hline
  Tight & \pbox{1.5\columnwidth}{
          Catania Science Gateway Framework~\cite{Ardizzone2012}, 
          Distributed application runtime environment (DARE~\cite{maddineni2012distributed}),
          DECIDE~\cite{ardizzone2012decide}$^{c,n}$, LONI Pipeline Environment$^n$~\cite{dinov2009efficient}
          }
  \\
  Service & \pbox{1.5\columnwidth}{
            Apache Airvata~\cite{marru2011apache}, e-BioInfra~\cite{shahand2015data}$^{w,n}$, HubZero with Pegasus~\cite{CPE:CPE3257}, MoSGrid~\cite{krüger2014mosgrid}$^w$, System in~\cite{wu2010accelerating}, Vine Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual Imaging Platform~\cite{GLAT-13}$^n$, WS-PGRADE/gUSE framework~\cite{Kacsuk2012}, 
            Science gateways in ~\cite{kacsuk2014science}$^w$
            } \\
  Sub-task & CBRAIN and PSOM~\cite{GLAT-16}$^n$, CBRAIN and FSL$^n$\\
  Pool & SHIWA pool~\cite{ROGE-13}\\
  Nested & \pbox{1.5\columnwidth}{SHIWA Simulation Platform (Coarse-Grained Interoperability~\cite{terstyanszky2014enabling})$^w$, HubZero with Pegasus (via hierarchical workflows)~\cite{Deelman201517} \todo{others}
  }\\
  Import & SHIWA Simulation Platform (Fine-Grained Interoperability)~\cite{plankensteiner-prodan-etal:2013}$^w$
\end{tabular}
\caption{Classification of science gateways based on the architecture used to integrate workflow engines. $^c$: based on the Catania Science Gateway Framework. $^w$: based on WS-PGRADE/gUSE. $^n$: used for neuroimaging.}
\label{table:system-classification}
\end{table*}

@@

This architecture is largely adopted, in systems such as Apache
Airvata~\cite{marru2011apache}, Vine
Toolkit~\cite{DBLP:journals/scpe/SzejnfeldDKKKKLPTWDNW10}, Virtual
Imaging Platform~\cite{GLAT-13}, the system
in~\cite{wu2010accelerating} and the WS-PGRADE/gUSE
framework~\cite{Kacsuk2012}. The integration between the
HubZero science gateway and the Pegasus workflow engine performed
in~\cite{CPE:CPE3257} also uses a service architecture where the
\texttt{d} interaction is implemented as a set of calls to Pegasus'
command-line tools (e.g. pegasus-status, pegasus-plan,
etc.). Figure~\ref{fig:vip-architecture} shows the architecture of the
Virtual Imaging Platform, where the MOTEUR workflow
engine~\cite{GLAT-08i} is invoked through a service.

@@

The pool model was implemented in the SHIWA pool~\cite{ROGE-13}
diagrammed in Figure~\ref{fig:shiwa-pool-architecture}. In this
Figure, interaction \texttt{d} of Figure~\subref*{archi:agent} is
implemented in 3 different types of calls to the pool: workflow
submission (\texttt{1} \& \texttt{2}), workflow status retrieval
(\texttt{a} and \texttt{b}), and workflow retrieval (\texttt{13} \&
\texttt{14}). In Figure~\ref{fig:shiwa-pool-architecture},
interaction \texttt{f} of Figure~\subref*{archi:agent} is implemented
through 2 types of calls: workflow instance retrieval (\texttt{4},
\texttt{5} and \texttt{6}), and workflow instance update (\texttt{11}
and \texttt{12}). Workflow instance retrieval is used by the agents to
fetch work from the pool. Workflow instance update is used by the
agents to update workflow statuses. In the SHIWA pool, agents can wrap
different types of workflow engines to execute workflows expressed
with different languages. Calls \texttt{0}, \texttt{7}, \texttt{8} and
\texttt{9} on Figure~\ref{fig:shiwa-pool-architecture} are used by
workflow engine plugins to declare their supported language and launch
engines, and by workflow engines to report their status to the
agent. These calls are specific to the SHIWA Pool implementation of
the \texttt{agent} component and therefore have no corresponding
representation in Figure~\subref*{archi:agent}.

@@

Nested workflows are abstract architectural patterns that can be
instantiated in the various architectures described previously. We
focus on instantiation with the service invocation model
(Figure~\subref*{archi:nested}-Right) as this is the most used
architecture. In such an instantiation, we assume that the parent and
child workflow engines are distinct pieces of software that require
different workflow services invoked by distinct \texttt{d}
interactions. If this is not the case, then workflow services can be
collapsed in a single one with a \texttt{d} interaction with
itself. An example of such interaction is the use of hierarchical
workflows in Pegasus~\cite{Deelman201517}.
Workflow engines communicate with infrastructures using
\texttt{b} and \texttt{c$_2$}. Science gateway transfers user data to
infrastructures using \texttt{c$_1$} interactions.

@@

See Figure~\subref*{archi:import}. This is an abstract model 
instantiated with the service invocation architecture for
consistency. Workflows are integrated in the science gateway through
format conversion from a native format to the science gateway
format. Workflow import is usually an offline process which is not
involved in workflow execution. Workflow import was implemented in the SHIWA
Simulation Platform through the IWIR language that provided a common
language for portability across grid workflow
systems~\cite{plankensteiner-prodan-etal:2013}.
